Accuracy and precision of zero-heat-flux temperature monitoring: a systematic review and meta- analysis \thanks{Grants or other notes about the article that should go on the front page should be placed here. General acknowledgments should be placed at the end of the article.}

### 2019-09-26 manuscript/manuscript.Rmd
results added - more work on figure needed

```{r meta-function, echo=FALSE}
meta <-  function(Te,V_T){
m <- length(Te)
wt_FE=1/V_T
T_FE <- sum(Te*wt_FE)/sum(wt_FE)
Q <- sum(wt_FE*(Te - T_FE)^2)
S1 <- sum(wt_FE)
S2 <- sum(wt_FE^2)
o2 <- (Q - (m - 1))/(S1 - S2/S1)
wt_RE <- 1/(V_T + o2)
T_RE <- sum(Te*wt_RE)/sum(wt_RE)
V_T_RE_mod <- 1/sum(wt_RE)
V_T_RE_rve <- (m/(m-1))*sum(wt_RE^2*(Te - T_RE)^2)/(sum(wt_RE))^2
c(m,T_RE,o2,V_T_RE_mod, V_T_RE_rve)
}
loa_maker <- function(bias,V_bias,logs2,V_logs2) {
bias_row=meta(bias, V_bias)
logs2_row=meta(logs2, V_logs2)
bias_mean <- bias_row[2]
sd2_est <- exp(logs2_row[2])
tau_est <- bias_row[3]
LOA_L <- bias_mean - 2*sqrt(sd2_est + tau_est)
LOA_U <- bias_mean + 2*sqrt(sd2_est + tau_est)
m <- bias_row[1]
tcrit <- qt(1-.05/2,m-1)
B1 <- sd2_est^2/(sd2_est + tau_est)
B2 <- tau_est^2/(sd2_est + tau_est)
wt <- 1/V_bias
S1 <- sum(wt)
S2 <- sum(wt^2)
S3 <- sum(wt^3)
A0 <- 2*(m-1)/(S1-S2/S1)^2
A1 <- 4/(S1 - S2/S1)
A2 <- 2*(S2-2*S3/S1+S2^2/S1^2)/(S1-S2/S1)^2
V_logT2 <- A0/tau_est^2 + A1/tau_est + A2
V_logT2 <- 2/sum((V_bias + tau_est)^(-2))
V_LOA_mod <- bias_row[4] + B1*logs2_row[4] + B2*V_logT2
V_LOA_rve <- bias_row[5] + B1*logs2_row[5] + B2*V_logT2
CI_L_mod <- LOA_L - tcrit*sqrt(V_LOA_mod)
CI_U_mod <- LOA_U + tcrit*sqrt(V_LOA_mod)
CI_L_rve <- LOA_L - tcrit*sqrt(V_LOA_rve)
CI_U_rve <- LOA_U + tcrit*sqrt(V_LOA_rve)
data.frame(m, bias_mean,sqrt(sd2_est), tau_est, LOA_L, LOA_U, CI_L_mod, CI_U_mod, CI_L_rve, CI_U_rve)
}
```

```{r load-data, include=FALSE}
library(tidyverse)
data <- readxl::read_xlsx(here::here("data","zhf_extracted.xlsx"))
```

@@

```{r format-result-function, include=FALSE}
format_results <- function(group){
out <- loa_maker(group$bias,group$V_bias, group$logs2, group$V_logs2)
out <- round(out, digits=1)
out <- out %>% 
  mutate(Participants = sum(group$n_count, na.rm=T)) %>% 
  mutate(Measurements = format(sum(group$N, na.rm = T), big.mark=",", scientific = FALSE)) %>% 
  mutate(Studies = length(unique(group$Study))) %>% 
  select(Studies, m, Participants, Measurements, bias_mean, tau_est, LOA_L, LOA_U, CI_L_rve, CI_U_rve) %>% 
  rename("Comparisons" = m, "Mean bias" = bias_mean, "Tau-squared" = tau_est, "Lower bound for 95% limit of agreement" = LOA_L, "Upper bound for 95% limits of agreement" = LOA_U, "Outer confidence interval for lower 95% limits of agreement" = CI_L_rve, "Outer confidence interval for upper 95% limits of agreement" = CI_U_rve)
out
}
```

@@

Core temperature monitoring is an important intervention within perioperative and intensive care settings. Thermoregulatory dysfunction is commonly associated with the induction of anesthesia and can lead to adverse outcomes including cardiac dysrhythmias, altered hemostasis and increased risk of surgical site infection [@frank1997perioperative; @kurz1996perioperative; @michelson1994reversible; @rohrer1992effect]. The pulmonary artery catheter is the reference standard for continuous core temperature monitoring$(ref - trouble finding a statement regarding this in the literature)$, however, the invasive nature of this method renders an increased risk of bloodstream infections and damage to surrounding tissue[@hadian2006evidence]. Although clinically accurate relative to pulmonary artery temperature measurements, surrogate measures of core temperature at esophageal, rectal, and bladder locations remain mildly invasive interventions$(ref)$.

Zero-heat-flux thermometry is an alternative non-invasive method that allows for continuous monitoring of core temperature. Originally developed in the 1970s, zero-heat-flux technology was recently implemented in the 3M SpotOn (3M, St Paul, MN), a single-use, disposable sensor $(ref)$. In practice, the zero-heat-flux sensor, such as the 3M SpotOn, is placed on the lateral surface of the forehead and initially warmed to equilibrate the temperature of the skin surface to the underlaying core tissues$(ref)$. Equipped with a thermal insulator, the zero-heat-flux sensor eliminates heat loss to the environment to allow for changes in core temperature to be directly reflected by a change in skin surface temperature $(ref)$.

The agreement between zero-heat-flux thermometers and core as well as other peripheral thermometers has been investigated in several studies conducted within clinical settings. Appraisal of these studies and synthesis of the results would aid clinicians in deciding the appropriate circumstances in which zero-heat-flux thermometers may be used. We aimed to determine if zero-heat-flux thermometers have clinically acceptable accuracy and precision relative to established core and peripheral temperature measurement devices. Accuracy is defined as the average difference between temperature measurements from the zero-heat-flux and comparator device and precision as the variance (standard deviation) in the differences.

@@

A systematic review was conducted in accordance with a predetermined protocol. The protocol was submitted to PROSPERO for registration but due to delays in processing we had started data extraction by the time it was reviewed. Therefore, the protocol did not meet the requirements for registration. A copy of the submitted protocol can be accessed [here](insert link). The primary comparison for this review was temperature measured from a zero-heat-flux thermometer versus a temperature measured from a core site, which we defined as arterial, esophaeal, bladder and rectal sites. Secondary comparisons were made between zero-heat-flux thermometers and temperatures taken at peripheral sites.

@@

Published studies were found by searching Medline and EMBASE from January 2000 to July 2019. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used [@mann2012should]. This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of zero-heat-flux thermometers. The search strategies used for each data base can be located in [file 1](). Selection of studies was undertaken independently by two reviewers using [Covidence](https://www.covidence.org).

@@

## Agreement between ZHF and core temperature measurements

@@

>>>>

```{r meta-function, echo=FALSE}
meta <-  function(Te,V_T){
m <- length(Te)
wt_FE=1/V_T
T_FE <- sum(Te*wt_FE)/sum(wt_FE)
Q <- sum(wt_FE*(Te - T_FE)^2)
S1 <- sum(wt_FE)
S2 <- sum(wt_FE^2)
o2 <- (Q - (m - 1))/(S1 - S2/S1)
wt_RE <- 1/(V_T + o2)
T_RE <- sum(Te*wt_RE)/sum(wt_RE)
V_T_RE_mod <- 1/sum(wt_RE)
V_T_RE_rve <- (m/(m-1))*sum(wt_RE^2*(Te - T_RE)^2)/(sum(wt_RE))^2
c(m,T_RE,o2,V_T_RE_mod, V_T_RE_rve)
}
loa_maker <- function(bias,V_bias,logs2,V_logs2) {
bias_row=meta(bias, V_bias)
logs2_row=meta(logs2, V_logs2)
bias_mean <- bias_row[2]
sd2_est <- exp(logs2_row[2])
tau_est <- bias_row[3]
LOA_L <- bias_mean - 2*sqrt(sd2_est + tau_est)
LOA_U <- bias_mean + 2*sqrt(sd2_est + tau_est)
m <- bias_row[1]
tcrit <- qt(1-.05/2,m-1)
B1 <- sd2_est^2/(sd2_est + tau_est)
B2 <- tau_est^2/(sd2_est + tau_est)
wt <- 1/V_bias
S1 <- sum(wt)
S2 <- sum(wt^2)
S3 <- sum(wt^3)
A0 <- 2*(m-1)/(S1-S2/S1)^2
A1 <- 4/(S1 - S2/S1)
A2 <- 2*(S2-2*S3/S1+S2^2/S1^2)/(S1-S2/S1)^2
V_logT2 <- A0/tau_est^2 + A1/tau_est + A2
V_logT2 <- 2/sum((V_bias + tau_est)^(-2))
V_LOA_mod <- bias_row[4] + B1*logs2_row[4] + B2*V_logT2
V_LOA_rve <- bias_row[5] + B1*logs2_row[5] + B2*V_logT2
CI_L_mod <- LOA_L - tcrit*sqrt(V_LOA_mod)
CI_U_mod <- LOA_U + tcrit*sqrt(V_LOA_mod)
CI_L_rve <- LOA_L - tcrit*sqrt(V_LOA_rve)
CI_U_rve <- LOA_U + tcrit*sqrt(V_LOA_rve)
data.frame(m, bias_mean, sd2_est, tau_est, LOA_L, LOA_U, CI_L_mod, CI_U_mod, CI_L_rve, CI_U_rve)
}
```

```{r load-data, include=FALSE}
data <- readxl::read_xlsx(here::here("data", "zhf_extracted.xlsx"))
```

@@

```{r format-result-function, include=FALSE}
format_results <- function(group){
out <- loa_maker(group$bias,group$V_bias, group$logs2, group$V_logs2)
out <- round(out, digits=1)
out <- out %>% 
  mutate(Participants = sum(group$n_count, na.rm=T)) %>% 
  mutate(Measurements = format(sum(group$N, na.rm = T), big.mark=",", scientific = FALSE)) %>% 
  mutate(Studies = length(unique(group$Study))) %>% 
  select(Studies, m, Participants, Measurements, bias_mean, sd2_est, tau_est, LOA_L, LOA_U, CI_L_rve, CI_U_rve) %>% 
  rename("Comparisons" = m)
out

# Better names for columns
# , "Mean bias" = bias_mean, "Tau-squared" = tau_est, "Lower bound for 95% limit of agreement" = LOA_L, "Upper bound for 95% limits of agreement" = LOA_U, "Outer confidence interval for lower 95% limits of agreement" = CI_L_rve, "Outer confidence interval for upper 95% limits of agreement" = CI_U_rve
}
```

@@

Core temperature monitoring is an important intervention within perioperative and intensive care settings. Thermoregulatory dysfunction is commonly associated with the induction of anesthesia and can lead to adverse outcomes including cardiac dysrhythmias, altered hemostasis and increased risk of surgical site infection [@frank1997perioperative; @kurz1996perioperative; @michelson1994reversible; @rohrer1992effect]. The pulmonary artery catheter is the reference standard for continuous core temperature monitoring, however, the invasive nature of this method renders an increased risk of bloodstream infections and damage to surrounding tissue[@hadian2006evidence]. Although clinically accurate relative to pulmonary artery temperature measurements, surrogate measures of core temperature at esophageal, rectal, and bladder locations remain mildly invasive interventions.

Zero-heat-flux thermometry is an alternative non-invasive method that allows for continuous monitoring of core temperature. Originally developed in the 1970s, zero-heat-flux technology was recently implemented in the 3M SpotOn temperature monitoring system (3M, St Paul, MN), as a single-use, disposable sensor. In practice, the zero-heat-flux sensor, such as the 3M SpotOn, is placed on the lateral surface of the forehead and isinitially warmed to equilibrate the temperature of the skin surface to the underlaying core tissues. Equipped with a thermal insulator, the zero-heat-flux sensor eliminates heat loss to the environment to allow for changes in core temperature to be directly reflected by a change in skin surface temperature [@eshraghi2014].

The agreement between zero-heat-flux thermometers and core as well as other peripheral thermometers has been investigated in multiple studies over the last five years. Appraisal of these studies and synthesis of the results would aid clinicians in deciding the appropriate circumstances in which zero-heat-flux thermometers may be used. We aimed to determine if zero-heat-flux thermometers have clinically acceptable accuracy and precision relative to established core and peripheral temperature measurement devices. Accuracy is defined as the average difference between temperature measurements from the zero-heat-flux and comparator device and precision as the variance (standard deviation) in the differences.

@@

A systematic review was conducted in accordance with a predetermined protocol. The protocol was submitted to PROSPERO for registration but due to delays in processing we had started data extraction by the time it was reviewed. Therefore, the protocol did not meet the requirements for registration. A copy of the submitted protocol can be accessed [here](https://github.com/awconway/zhf-review). The primary comparison for this review was temperature measured from a zero-heat-flux thermometer versus temperature measured from a core site, which we defined as temperature measured at either an arterial, esophaeal, bladder or rectal site. Secondary comparisons were made between zero-heat-flux thermometers and temperatures taken at peripheral sites.

@@

Published studies were found by searching Medline and EMBASE from January 2000 to July 2019. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used [@mann2012should]. This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of zero-heat-flux thermometers. The search strategies used for each data base can be located [here](shiny app). Selection of studies was undertaken independently by two reviewers using [Covidence](https://www.covidence.org).

@@

## Primary comparison: zero-heat-flux thermometer versus core thermometers

The pooled estimate for the mean bias between zero-heat-flux and core temperature measurements was `r primary$bias_mean`$^\circ C$. However, the variation in differences between studies was large, as displayed by the density plot in Figure 2. Consequently, the population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`$^\circ C$ to `r primary$CI_U_rve`$^\circ C$ (`r primary$Measurements` measurements; `r primary$Participants` participants; `r primary$Studies` studies). The quality of evidence for the primary comparison was downgraded to low quality due to concerns about study limitations and inconsistency. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QADAS-2 were similar to the primary analysis (`r low_risk$Measurements` measurements; `r low_risk$Participants` participants; `r low_risk$Studies` stsudies). The mean bias was again `r low_risk$bias_mean`$^\circ C$ with population agreements spanning from `r low_risk$CI_L_rve`$^\circ C$ to `r low_risk$CI_U_rve`$^\circ C$.

```{r fig-2, fig.align="center", fig.cap="Comparisons between core and zero-heat-flux thermometers within and across studies. Blue curves are distributions of the differences between measurements from zero-heat-flux (ZHF) sensors and core temperature measurements in individual studies. The red curve is the distribution of the pooled estimate.", echo=FALSE, fig.width=5.1, fig.path=""}

    bias=data_core$bias
    s2_unb = data_core$s2
    pooled_bias = primary$bias_mean
    pooled_sd = primary$sd2_est
    pooled_tau2 = primary$tau_est
    pooled_sd = sqrt(pooled_sd^2 + pooled_tau2)

    LOA_l = primary$LOA_L
    LOA_u = primary$LOA_U
    LOA_l_CI = primary$CI_L_rve
    LOA_u_CI = primary$CI_U_rve

g <- ggplot(data.frame(x=seq(-20,20,length=200)), aes(x=x)) + 
      stat_function(fun=dnorm, args = list(bias[1], sd=sqrt(s2_unb[1])), colour = "cornflowerblue", size=0.6, alpha = 0.5)

    for (i in 2:length(bias)){
      g <- g + stat_function(fun=dnorm, args = list(bias[i], sd=sqrt(s2_unb[i])), color = "cornflowerblue", size = 0.6, alpha = 0.5)
    }

    g <- g + stat_function(fun=dnorm, args = list(pooled_bias, pooled_sd), colour ="lightcoral", alpha = 0.05) + 
      stat_function(fun=dnorm, args = list(pooled_bias, pooled_sd), colour = NA, geom="area", fill="lightcoral", alpha = 0.6) + 
      scale_x_continuous(limits = c(-3.5,3.5)) + labs(x = expression('Difference between core and ZHF temperature measurements in'*~degree*C*''))+
      scale_y_continuous(name = "Density \n", limits = c(0,2.5)) +  
      #labs(title = "\nOuter confidence intervals for pooled limits of agreement\n\n  Pooled limits of agreement")+
      theme(plot.title = element_text(hjust = 0.5, margin = margin(t=10, b=-32), size=10),
            axis.text=element_text(hjust = 0.5, size=10),                                                                        
            axis.title=element_text(hjust = 0.5, size=10),
            axis.line=element_line(colour="black", size=0.2),
            panel.background = element_blank(),
            axis.ticks = element_blank(), 
            plot.caption = element_text(hjust = 0)) #+ 
      #CI lines that extend to arrow tips
      # geom_segment(aes(x=LOA_u, y=0, xend=LOA_u, yend=2.25), size = 0.3, linetype="dashed")+ 
      # geom_segment(aes(x=LOA_l, y=0, xend=LOA_l, yend=2.25), size = 0.3, linetype="dashed")+
      # geom_segment(aes(x=LOA_u_CI, y=0, xend=LOA_u_CI, yend=2.48), size = 0.3)+
      # geom_segment(aes(x=LOA_l_CI, y=0, xend=LOA_l_CI, yend=2.48), size = 0.3)#+
      # 
      # geom_segment(aes(x=0, y=2.48, xend=LOA_u_CI, yend=2.48), size = 0.4, arrow = arrow(length = unit(0.03, "npc")))+ #R arrow high
      # geom_segment(aes(x=0, y=2.48, xend=LOA_l_CI, yend=2.48), size = 0.4, arrow = arrow(length = unit(0.03, "npc")))+ #L arrow high
      # geom_segment(aes(x=0, y=2.25, xend=LOA_u, yend=2.25), size = 0.4,  arrow = arrow(length = unit(0.03, "npc")))+ #R arrow low
      # geom_segment(aes(x=0, y=2.25, xend=LOA_l, yend=2.25), size = 0.4,  arrow = arrow(length = unit(0.03, "npc"))) #L arrow low
g
```

We conducted two subgroup analyses for the primary comparison according to the clinical setting in which the study was conducted. In the subset of studies conducted in the ICU, the mean bias was `r ICU$bias_mean` $^\circ C$ with population limits of agreement between `r ICU$CI_L_rve` $^\circ C$ and `r ICU$CI_U_rve` $^\circ C$. Whereas in the subset of studies that evaluated the use of zero-heat-flux thermometers during surgery found a mean bias of `r OT$bias_mean` $^\circ C$ with population limits of agreement from `r OT$CI_L_rve` $^\circ C$ to `r OT$CI_U_rve` $^\circ C$. The quality of evidence for these subgroup analyses was similarly rated as low quality due to concerns about study limitations and inconsistency.

## Secondary comparisons: zero-heat-flux thermometer versus peripheral thermometers

Zero-heat-flux thermometers were compared with sublingual thermometers and nasopharyngeal thermometers in the studies included in this review. The mean bias between zero-heat-flux and sublingual temperature measurements in meta-analysis of results from `r SL$Studies` studies was `r SL$bias_mean` $^\circ C$. Due to the limited number of studies and measurements, population limits of agreement were extremely wide, spanning from `r SL$CI_L_rve` $^\circ C$ to `r SL$CI_U_rve` $^\circ C$. The quality of evidence for this comparison was rated as very low quality due to serious concerns about imprecision.

The mean bias between zero-heat-flux and nasopharyngeal temperature measurements in meta-analysis of results from `r NPA$Studies` studies was `r NPA$bias_mean` $^\circ C$. Population limits of agreement were `r NPA$CI_L_rve` $^\circ C$ to `r NPA$CI_U_rve` $^\circ C$. We downgraded the quality of evidence to low, again due to concerns about inconsistency and study limitations.

@@

Results from this systematic review have important implications for practice. Clinicians should consider the potential that a temperature measurement from a zero-heat-flux thermometer could be as much as one $^\circ C$ higher or lower than the true core temperature. This is similar to results from previous meta-analyses that compared other peripheral thermometers with core temperature measurements. However, these previous meta-analyses used statistical approaches that did not incorporate the magnitude of heteroogeneity in results between studies or sampling error. As such, it is possible that the zero-heat-flux thermometer is still more precise than these other peripheral thermometers.

Whether or not teh zero-heat-flux thermometer is sufficiently accurate to be used in place of nasopharyngeal thermometers is unclear.

### 2019-09-26 manuscript/manuscript.tex
results added - more work on figure needed

    \subclass{
                    MSC code 1 \and
                    MSC code 2 \and
            }

@@

Core temperature monitoring is an important intervention within
perioperative and intensive care settings. Thermoregulatory dysfunction
is commonly associated with the induction of anesthesia and can lead to
adverse outcomes including cardiac dysrhythmias, altered hemostasis and
increased risk of surgical site infection (Frank et al. 1997; Kurz,
Sessler, and Lenhardt 1996; Michelson et al. 1994; Rohrer and Natale
1992). The pulmonary artery catheter is the reference standard for
continuous core temperature
monitoring\((ref - trouble finding a statement regarding this in the literature)\),
however, the invasive nature of this method renders an increased risk of
bloodstream infections and damage to surrounding tissue(Hadian and
Pinsky 2006). Although clinically accurate relative to pulmonary artery
temperature measurements, surrogate measures of core temperature at
esophageal, rectal, and bladder locations remain mildly invasive
interventions\((ref)\).

Zero-heat-flux thermometry is an alternative non-invasive method that
allows for continuous monitoring of core temperature. Originally
developed in the 1970s, zero-heat-flux technology was recently
implemented in the 3M SpotOn (3M, St Paul, MN), a single-use, disposable
sensor \((ref)\). In practice, the zero-heat-flux sensor, such as the 3M
SpotOn, is placed on the lateral surface of the forehead and initially
warmed to equilibrate the temperature of the skin surface to the
underlaying core tissues\((ref)\). Equipped with a thermal insulator,
the zero-heat-flux sensor eliminates heat loss to the environment to
allow for changes in core temperature to be directly reflected by a
change in skin surface temperature \((ref)\).

The agreement between zero-heat-flux thermometers and core as well as
other peripheral thermometers has been investigated in several studies
conducted within clinical settings. Appraisal of these studies and
synthesis of the results would aid clinicians in deciding the
appropriate circumstances in which zero-heat-flux thermometers may be
used. We aimed to determine if zero-heat-flux thermometers have
clinically acceptable accuracy and precision relative to established
core and peripheral temperature measurement devices. Accuracy is defined
as the average difference between temperature measurements from the
zero-heat-flux and comparator device and precision as the variance
(standard deviation) in the differences.

@@

A systematic review was conducted in accordance with a predetermined
protocol. The protocol was submitted to PROSPERO for registration but
due to delays in processing we had started data extraction by the time
it was reviewed. Therefore, the protocol did not meet the requirements
for registration. A copy of the submitted protocol can be accessed
\href{insert\%20link}{here}. The primary comparison for this review was
temperature measured from a zero-heat-flux thermometer versus a
temperature measured from a core site, which we defined as arterial,
esophaeal, bladder and rectal sites. Secondary comparisons were made
between zero-heat-flux thermometers and temperatures taken at peripheral
sites.

@@

Published studies were found by searching Medline and EMBASE from
January 2000 to July 2019. The Cochrane-recommended search strategy
combining terms for the `target condition' and `index test' was used
(Mann and Gilbody 2012). This search strategy is an efficient approach
for systematic reviews of diagnostic test accuracy studies.(Preston et
al. 2015) We also conducted forward citation searching, by using Google
Scholar to search the citations of the first article published on the
accuracy of zero-heat-flux thermometers. The search strategies used for
each data base can be located in \href{}{file 1}. Selection of studies
was undertaken independently by two reviewers using
\href{https://www.covidence.org}{Covidence}.

@@

\hypertarget{agreement-between-zhf-and-core-temperature-measurements}{%
\subsection{Agreement between ZHF and core temperature
measurements}\label{agreement-between-zhf-and-core-temperature-measurements}}

@@

@@

>>>>

@@

Core temperature monitoring is an important intervention within
perioperative and intensive care settings. Thermoregulatory dysfunction
is commonly associated with the induction of anesthesia and can lead to
adverse outcomes including cardiac dysrhythmias, altered hemostasis and
increased risk of surgical site infection (Frank et al. 1997; Kurz,
Sessler, and Lenhardt 1996; Michelson et al. 1994; Rohrer and Natale
1992). The pulmonary artery catheter is the reference standard for
continuous core temperature monitoring, however, the invasive nature of
this method renders an increased risk of bloodstream infections and
damage to surrounding tissue(Hadian and Pinsky 2006). Although
clinically accurate relative to pulmonary artery temperature
measurements, surrogate measures of core temperature at esophageal,
rectal, and bladder locations remain mildly invasive interventions.

Zero-heat-flux thermometry is an alternative non-invasive method that
allows for continuous monitoring of core temperature. Originally
developed in the 1970s, zero-heat-flux technology was recently
implemented in the 3M SpotOn temperature monitoring system (3M, St Paul,
MN), as a single-use, disposable sensor. In practice, the zero-heat-flux
sensor, such as the 3M SpotOn, is placed on the lateral surface of the
forehead and isinitially warmed to equilibrate the temperature of the
skin surface to the underlaying core tissues. Equipped with a thermal
insulator, the zero-heat-flux sensor eliminates heat loss to the
environment to allow for changes in core temperature to be directly
reflected by a change in skin surface temperature (Eshraghi et al.
2014).

The agreement between zero-heat-flux thermometers and core as well as
other peripheral thermometers has been investigated in multiple studies
over the last five years. Appraisal of these studies and synthesis of
the results would aid clinicians in deciding the appropriate
circumstances in which zero-heat-flux thermometers may be used. We aimed
to determine if zero-heat-flux thermometers have clinically acceptable
accuracy and precision relative to established core and peripheral
temperature measurement devices. Accuracy is defined as the average
difference between temperature measurements from the zero-heat-flux and
comparator device and precision as the variance (standard deviation) in
the differences.

@@

A systematic review was conducted in accordance with a predetermined
protocol. The protocol was submitted to PROSPERO for registration but
due to delays in processing we had started data extraction by the time
it was reviewed. Therefore, the protocol did not meet the requirements
for registration. A copy of the submitted protocol can be accessed
\href{https://github.com/awconway/zhf-review}{here}. The primary
comparison for this review was temperature measured from a
zero-heat-flux thermometer versus temperature measured from a core site,
which we defined as temperature measured at either an arterial,
esophaeal, bladder or rectal site. Secondary comparisons were made
between zero-heat-flux thermometers and temperatures taken at peripheral
sites.

@@

Published studies were found by searching Medline and EMBASE from
January 2000 to July 2019. The Cochrane-recommended search strategy
combining terms for the `target condition' and `index test' was used
(Mann and Gilbody 2012). This search strategy is an efficient approach
for systematic reviews of diagnostic test accuracy studies.(Preston et
al. 2015) We also conducted forward citation searching, by using Google
Scholar to search the citations of the first article published on the
accuracy of zero-heat-flux thermometers. The search strategies used for
each data base can be located \href{shiny\%20app}{here}. Selection of
studies was undertaken independently by two reviewers using
\href{https://www.covidence.org}{Covidence}.

@@

\hypertarget{primary-comparison-zero-heat-flux-thermometer-versus-core-thermometers}{%
\subsection{Primary comparison: zero-heat-flux thermometer versus core
thermometers}\label{primary-comparison-zero-heat-flux-thermometer-versus-core-thermometers}}

The pooled estimate for the mean bias between zero-heat-flux and core
temperature measurements was 0\(^\circ C\). However, the variation in
differences between studies was large, as displayed by the density plot
in Figure 2. Consequently, the population limits of agreement, which
take into consideration the between-study heterogeneity and sampling
error, were wide, spanning from -1\(^\circ C\) to 1.1\(^\circ C\)
(179,821 measurements; 576 participants; 13 studies). The quality of
evidence for the primary comparison was downgraded to low quality due to
concerns about study limitations and inconsistency. Population limits of
agreement for the sensitivity analysis restricted to studies rated as
having low risk of bias across all the domains of the QADAS-2 were
similar to the primary analysis (104,294 measurements; 273 participants;
5 stsudies). The mean bias was again 0\(^\circ C\) with population
agreements spanning from -1\(^\circ C\) to 1.1\(^\circ C\).

{\centering \includegraphics{fig-2-1}

\caption{Comparisons between core and zero-heat-flux thermometers within and across studies. Blue curves are distributions of the differences between measurements from zero-heat-flux (ZHF) sensors and core temperature measurements in individual studies. The red curve is the distribution of the pooled estimate.}\label{fig:fig-2}
\end{figure}

We conducted two subgroup analyses for the primary comparison according
to the clinical setting in which the study was conducted. In the subset
of studies conducted in the ICU, the mean bias was 0 \(^\circ C\) with
population limits of agreement between -1.5 \(^\circ C\) and 1.4
\(^\circ C\). Whereas in the subset of studies that evaluated the use of
zero-heat-flux thermometers during surgery found a mean bias of 0
\(^\circ C\) with population limits of agreement from -1.1 \(^\circ C\)
to 1.1 \(^\circ C\). The quality of evidence for these subgroup analyses
was similarly rated as low quality due to concerns about study
limitations and inconsistency.

\hypertarget{secondary-comparisons-zero-heat-flux-thermometer-versus-peripheral-thermometers}{%
\subsection{Secondary comparisons: zero-heat-flux thermometer versus
peripheral
thermometers}\label{secondary-comparisons-zero-heat-flux-thermometer-versus-peripheral-thermometers}}

Zero-heat-flux thermometers were compared with sublingual thermometers
and nasopharyngeal thermometers in the studies included in this review.
The mean bias between zero-heat-flux and sublingual temperature
measurements in meta-analysis of results from 2 studies was -0.2
\(^\circ C\). Due to the limited number of studies and measurements,
population limits of agreement were extremely wide, spanning from -13.3
\(^\circ C\) to 12.8 \(^\circ C\). The quality of evidence for this
comparison was rated as very low quality due to serious concerns about
imprecision.

The mean bias between zero-heat-flux and nasopharyngeal temperature
measurements in meta-analysis of results from 4 studies was 0
\(^\circ C\). Population limits of agreement were -1 \(^\circ C\) to 1
\(^\circ C\). We downgraded the quality of evidence to low, again due to
concerns about inconsistency and study limitations.

@@

Results from this systematic review have important implications for
practice. Clinicians should consider the potential that a temperature
measurement from a zero-heat-flux thermometer could be as much as one
\(^\circ C\) higher or lower than the true core temperature. This is
similar to results from previous meta-analyses that compared other
peripheral thermometers with core temperature measurements. However,
these previous meta-analyses used statistical approaches that did not
incorporate the magnitude of heteroogeneity in results between studies
or sampling error. As such, it is possible that the zero-heat-flux
thermometer is still more precise than these other peripheral
thermometers.

Whether or not teh zero-heat-flux thermometer is sufficiently accurate
to be used in place of nasopharyngeal thermometers is unclear.

@@

\leavevmode\hypertarget{ref-eshraghi2014}{}%
Eshraghi, Yashar, Vivian Nasr, Ivan Parra-Sanchez, Albert Van Duren,
Mark Botham, Thomas Santoscoy, and Daniel I Sessler. 2014. ``An
Evaluation of a Zero-Heat-Flux Cutaneous Thermometer in Cardiac Surgical
Patients.'' \emph{Anesthesia \& Analgesia} 119 (3). LWW: 543--49.
